outline!

### 2016-05-24 paper.tex
paper: add some code descriptions
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

Listing~\ref{lab} shows code.

\begin{lstlisting}[caption={Epoch Guard}, label=lst:eg]
bloom :epoch_guard do
  invalid <= (op * epoch).pairs{|o,e| o.epoch <= e.epoch}
  ret <= invalid{|o| [o.type, o.pos, o.epoch, 'stale']}
  valid <= op.notin(invalid)
end
\end{lstlisting}

>>>>

\begin{lstlisting}[caption={Epoch Guard}, label=lst:eg]
state do
  table :epoch, [:epoch]
  table :log, [:pos] => [:state, :data]

  interface input,  :op,   [:type, :pos, :epoch] => [:data]
  interface output, :ret,  [:type, :pos, :epoch] => [:retval]

  scratch :write_op, [:type, :pos, :epoch] => [:data]
  scratch :read_op,  [:type, :pos, :epoch] => [:data]
  scratch :trim_op,  [:type, :pos, :epoch] => [:data]
  scratch :fill_op,  [:type, :pos, :epoch] => [:data]
  scratch :seal_op,  [:type, :pos, :epoch] => [:data]

  # op did or did not pass the epoch guard
  scratch :valid_op,   [:type, :pos, :epoch] => [:data]
  scratch :invalid_op, [:type, :pos, :epoch] => [:data]

  # op's position was or was not found in the log
  scratch :found_op,    [:type, :pos, :epoch] => [:data]
  scratch :notfound_op, [:type, :pos, :epoch] => [:data]
end

bloom do
  # op's position found in log
  found_op <= (valid_op * log).lefts(pos => pos)
  notfound_op <= valid_op.notin(found_op)

  # demux on operation type
  write_op <= valid_op { |o| o if o.type == 'write' }
  read_op  <= valid_op { |o| o if o.type == 'read'  }
  fill_op  <= valid_op { |o| o if o.type == 'fill'  }
  trim_op  <= valid_op { |o| o if o.type == 'trim'  }
  seal_op  <= valid_op { |o| o if o.type == 'seal'  }
end
\end{lstlisting}

Listing~\ref{lst:eg} shows the epoch guard that is applied to all operations.
The guard rejects requests that are tagged with old epoch values, ensuring that
a client generating a request has an up-to-date view of the system.  First the
\texttt{invalid\_op} collection is defined to include the current operation if
its epoch value is no larger than the stored epoch value. Next the
\texttt{valid\_op} collection is defined to be the inverse of
\texttt{invalid\_op} and is a helper used to refine other operations later in
the dataflow. Finally we handle the case for all operations tagged with an
out-of-date epoch by merging the \texttt{invalid\_op} set into the output
\texttt{ret} collection.

\begin{lstlisting}[caption={Epoch Guard}, label=lst:eg]
bloom :epoch_guard do
  invalid_op <= (op * epoch).pairs{|o,e| o.epoch <= e.epoch}
  valid_op <= op.notin(invalid_op)
  ret <= invalid_op{|o| [o.type, o.pos, o.epoch, 'stale']}
end
\end{lstlisting}

The process of sealing an object requires installing a new epoch value and
returning the current maximum position written. Listing~\ref{lst:seal} 
implements the seal interface by first removing the current epoch value and
replacing it with the epoch value contained in the operation. Next an aggregate
is computed over the log to find the maximum position written, and this value
is returned.

\begin{lstlisting}[caption={Seal}, label=lst:seal]
bloom :seal do
  epoch <- (seal_op * epoch).rights
  epoch <+ seal_op { |o| [o.epoch] }
  temp :maxpos <= log.group([], max(pos))
  ret <= (seal_op * maxpos).pairs do |o, m|
    [o.type, nil, o.epoch, m.content]
  end
end
\end{lstlisting}

Trimming a log entry always succeeds. In Listing~\ref{lst:trim} The
\texttt{<+-} operator simulatenously removes the log entry with the given 
position and replaces it with an entry with its state set to \emph{trimmed}.

\begin{lstlisting}[caption={Trim}, label=lst:trim]
bloom :trim do
  log <+- trim_op{|op| [op.pos, 'trimmed']}
  ret <=  trim_op{|op| [op.type, op.pos, op.epoch, 'ok']}
end
\end{lstlisting}

The write and fill interfaces are implemented similarily, and are show in
Listing~\ref{lst:write} and Listing~\ref{lst:fill}, respectively.  A valid
write collection is created if the operation position is not found in the log.
A valid write operation is then merged into log, otherwise a read only error is
returend indicating that hte log position was already written to.
The fill operation is identical except a the fill state is set on the log entry.

\begin{lstlisting}[caption={Write}, label=lst:write]
bloom :write do
  temp :valid_write <= write_op.notin(found_op)
  log <+ valid_write { |o| [o.pos, 'valid', o.data] }
  ret <= valid_write { |o| [o.type, o.pos, o.epoch, 'ok'] }
  ret <= write_op.notin(valid_write) { |o|
    [o.type, o.pos, o.epoch, 'read-only']
  }
end
\end{lstlisting}

\begin{lstlisting}[caption={Fill}, label=lst:fill]
bloom :fill do
  temp :valid_fill <= fill_op.notin(found_op)
  log <+ valid_fill { |o| [o.pos, 'fill'] }
  ret <= valid_fill { |o| [o.type, o.pos, o.epoch, 'ok'] }
  ret <= fill_op.notin(valid_fill) { |o|
    [o.type, o.pos, o.epoch, 'read-only']
  }
end
\end{lstlisting}

The read interface is shown in Listing~\ref{lst:read}, and structured similar
to the write interface. First we create a collection containing a valid read
operation that is in the log and does not have the filled or trimmed state set.
The data is returend in the case of a valid read op, otherwise an error is
returend through the output interface.

\begin{lstlisting}[caption={Read}, label=lst:read]
bloom :read do
  temp :valid_read <= (read_op * log).pairs(pos => pos) { |o, l|
    [o.type, o.pos, o.epoch, l.data] unless
      ['filled', 'trimmed'].include?(l.state)
  }
  ret <= valid_read { |e| [e.type, e.pos, e.epoch, e.data] }
  ret <= read_op.notin(valid_read, type => type) do |o|
    [o.type, o.pos, o.epoch, 'invalid']
  end
end
\end{lstlisting}
